Equalizer Programming Guide

### 2008-05-10 Developer/ProgrammingGuide/paper.tex
Website: Start file format section in Programming Guide
[ ] May break build
[ ] Breaks existing applications
[ ] Bugfix
[ ] New Feature
[ ] Cleanup
[ ] Optimization
[x] Documentation

\usepackage{tabularx,graphicx,a4,listings,wrapfig,subfigure,textcomp}

@@

\subsection*{Previous Page}

The images on the front page show: a terrain rendering application on a
six-node display wall [top left], the \textsf{eqPly} polygonal renderer
in a three-sided CAVE [middle], two volume rendering results from
eVolve\footnote{Data sets courtesy of General Electric, USA and AVS,
  USA} [bottom left and middle] and a six-node sort-last database
decomposition with parallel direct-send recomposition\footnote{Data set
  courtesy of Stanford University Computer Graphics Laboratory} [bottom
right].

\clearpage
\thispagestyle{empty}
\tableofcontents
\vfill{\center\begin{tabularx}{\textwidth}{|l|l|X|}
    \hline
    \bf Rev & \bf Date     & \bf Changes \\
    \hline
    1.0     & Oct 28, 2007 & Initial Version for Equalizer 0.4\\
    1.2     & Apr 15, 2008 & Version for Equalizer 0.5\\
    1.3.1   & May 02, 2008 & Add statistics, improved images based on 1.2 print\\
    1.3.2   & May 05, 2008 & Add network distribution on eqPly model data\\
    \hline
  \end{tabularx}}
\thispagestyle{empty}
\clearpage

@@

The resource usage is configured using a compound tree, which is a
hierarchical representation of the rendering decomposition and
recomposition across the resources. It is explained in \sref{compounds}.

@@

\subsection{\label{compounds}Compounds}

@@

>>>>

\usepackage{tabularx,longtable,graphicx,a4,listings,wrapfig,subfigure,textcomp}

@@

\subsection*{Front Page}

The images on the front page show the following Equalizer applications:
a terrain rendering application on a six-node display wall [top left], a
flow visualization application for climate research\footnote{Image
  courtesy of Computer Graphics and Multimedia Systems, University of
  Siegen}, the \textsf{eVolve} volume renderer\footnote{Data sets
  courtesy of General Electric, USA and AVS, USA} [bottom left] the
\textsf{eqPly} polygonal renderer in a three-sided CAVE [bottom center],
and a six-node sort-last database decomposition with parallel
direct-send recomposition\footnote{Data set courtesy of Stanford
  University Computer Graphics Laboratory} [bottom right].

\clearpage
\thispagestyle{empty}
\tableofcontents
\vfill{\center\begin{tabularx}{\textwidth}{|l|l|X|}
    \hline
    \bf Rev & \bf Date     & \bf Changes \\
    \hline
    1.0     & Oct 28, 2007 & Initial Version for Equalizer 0.4\\
    1.2     & Apr 15, 2008 & Version for Equalizer 0.5\\
    1.3.1   & May 02, 2008 & Add statistics, improved images based on 1.2 print\\
    1.3.2   & May 05, 2008 & Add network distribution on eqPly model data\\
    1.3.3   & May 10, 2008 & File Format Appendix [incomplete]\\
    \hline
  \end{tabularx}}
\thispagestyle{empty}
\clearpage

@@

The resource usage is configured using a compound tree, which is a
hierarchical representation of the rendering decomposition and
recomposition across the resources. It is explained in \sref{sCompounds}.

@@

\subsection{\label{sCompounds}Compounds}

@@

\newpage
\appendix
\section{File Format}

The currently used file format is the one-to-one representation of the
server's internal data structures. Its purpose is intermediate, that
is, it will gradually be replaced by automatic resource dedection and
configuration. However the core scalability engine will always use a
similar structure as currently exposed by the file format.

The file format represents an ascii deserialization of the
server. Streaming an \textsf{eqs::Server} to an \textsf{eqBase::Log}
ostream produces a valid configuration file. Likewise, loading a
configuration file produces an \textsf{eqs::Server}.

The file format uses the same syntactical structure as VRML. If your
editor supports syntax highlighting and formatting for VRML, you can use
this mode for \textsf{.eqc} files.

The configuration file consist of an optional global section and a
server configuration. The global section defines default values for
various attributes. The server section represents an
\textsf{eqs::Server}.

\subsection{Global Section}

The global section defines default values for attributes used by the
individual entities of the server section. The naming convention for
attributes is:

{\footnotesize\begin{lstlisting}
EQ_<ENTITY>_<DATATYPE>ATTR_<ATTR_NAME>
\end{lstlisting}}

The entity is the capitalized name of the corresponding section later in
the configuration file: connection, config, pipe, window, channel or
compound. The connection is used by the server and nodes.

The datatype is one capital letter for the type of the attribute's
value: \textsf{S} for strings, \textsf{C} for a character, \textsf{I}
for an integer and \textsf{F} for floating-point values. Enumation
values are handled as integers. Strings have always to be surrounded by
double quotes '\textsf{''}'. A character has to surrounded by single
quotes '\textsf{'}'.

The attribute name is the capitalized name of the entities attribute, as
discussed in the following sections.

Global attribute values have useful default parameters, which can be
overriden with an environment variable of the same name. For enumeration
values the corresponding integer value has to be used. The global values
in the config file override these defaults, and are in turn overriden by
the corresponding attributes sections of the associated entities.

The globals section starts with the token \textsf{global} and an open
curly brace '\textsf{\{}', and is terminated with a closing curly brace
'\textsf{\}}'. Within the braces, globals are set using the attribute's
name followed by its value. The following attributes are available:

\begin{center}
\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_CONNECTION\_SATTR\_HOSTNAME\\
  \textbf{Value} & string\\
  \textbf{Default} & ''localhost''\\
  \textbf{Description} & The hostname or IP address used to connect the
  server or node. When used for the server, the listening port of the
  server is bound to this address. When used for a node, the server
  first tries to connect to the render client node using this hostname,
  and then tries to launch the render client executable on this host.\\
  \hline
\end{tabularx}\\[\smallskipamount]

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_CONNECTION\_SATTR\_LAUNCH\_COMMAND\\
  \textbf{Value} & string\\
  \textbf{Default} & ssh -n \%h \%c \textgreater\& \%h.\%n.log [POSIX]\\
                   & ssh -n \%h \%c [WIN32]\\
  \textbf{Description} & The command used by the server to launch nodes
  which could not be connected. The launch command is executed from a
  process forked of the server process. The \% tokens are replaced by the
  server at runtime with concrete data: \%h is replaced by the hostname,
  \%c by the command to launch, including command line arguments and \%n
  by a node-unique identifier. Each command line argument are surrounded
  by launch command quotes.\\
  \hline
\end{tabularx}\\[\smallskipamount]

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_CONNECTION\_CATTR\_LAUNCH\_COMMAND\_QUOTE\\
  \textbf{Value} & character\\
  \textbf{Default} & ' [POSIX]\\
                   & '' [WIN32]\\
  \textbf{Description} & The server uses command line arguments to
  launch render client nodes correctly. This arguments might contain
  white spaces, and therefore have to be surrounded by quotes to
  identify their borders. Certain launch commands or shells use
  different conventions to separate command line arguments.\\
  \hline
\end{tabularx}\\[\smallskipamount]

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_CONNECTION\_IATTR\_TYPE\\
  \textbf{Value} & TCPIP \textbar \ SDP\\
  \textbf{Default} & TCPIP\\
  \textbf{Description} & The protocol for connections. SDP
  programmatically selects the socket direct protocol (AF\_INET\_SDP)
  provided by most InfiniBand protocol stacks.\\
  \hline
\end{tabularx}\\[\smallskipamount]

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_CONNECTION\_IATTR\_TCPIP\_PORT\\
  \textbf{Value} & unsigned\\
  \textbf{Default} & 0\\
  \textbf{Description} & The listening port used by the server or
  node. For nodes, the port can be used to contact pre-started, resident
  render client nodes or to use a specific port for the node. If 0 is
  specified, a random port is chosen. Note that a server with no
  connections automatically creates a default connection using the
  server's default port.\\
  \hline
\end{tabularx}\\[\smallskipamount]

\begin{tabularx}{\textwidth}{|l|X|}
\hline
\textbf{Name} & EQ\_CONNECTION\_IATTR\_LAUNCH\_TIMEOUT\\
\textbf{Value} & unsigned\\
\textbf{Default} & \\
\textbf{Description} & \\
\hline
\end{tabularx}\\[\smallskipamount]

\begin{tabularx}{\textwidth}{|l|X|}
\hline
\textbf{Name} & EQ\_CONFIG\_FATTR\_EYE\_BASE\\
\textbf{Value} & float\\
\textbf{Default} & \\
\textbf{Description} & \\
\hline
\end{tabularx}\\[\smallskipamount]

\begin{tabularx}{\textwidth}{|l|X|}
\hline
\textbf{Name} & EQ\_PIPE\_IATTR\_HINT\_THREAD\\
\textbf{Value} & OFF \textbar \ ON\\
\textbf{Default} & \\
\textbf{Description} & \\
\hline
\end{tabularx}\\[\smallskipamount]

\begin{tabularx}{\textwidth}{|l|X|}
\hline
\textbf{Name} & EQ\_WINDOW\_IATTR\_HINT\_STEREO\\
\textbf{Value} & OFF \textbar \ ON \textbar \ AUTO\\
\textbf{Default} & \\
\textbf{Description} & \\
\hline
\end{tabularx}\\[\smallskipamount]

\begin{tabularx}{\textwidth}{|l|X|}
\hline
\textbf{Name} & EQ\_WINDOW\_IATTR\_HINT\_DOUBLEBUFFER\\
\textbf{Value} & OFF \textbar \ ON \textbar \ AUTO\\
\textbf{Default} & \\
\textbf{Description} & \\
\hline
\end{tabularx}\\[\smallskipamount]

\begin{tabularx}{\textwidth}{|l|X|}
\hline
\textbf{Name} & EQ\_WINDOW\_IATTR\_HINT\_DECORATION\\
\textbf{Value} & OFF \textbar \ ON\\
\textbf{Default} & \\
\textbf{Description} & \\
\hline
\end{tabularx}\\[\smallskipamount]

\begin{tabularx}{\textwidth}{|l|X|}
\hline
\textbf{Name} & EQ\_WINDOW\_IATTR\_HINT\_FULLSCREEN\\
\textbf{Value} & OFF \textbar \ ON\\
\textbf{Default} & \\
\textbf{Description} & \\
\hline
\end{tabularx}\\[\smallskipamount]

\begin{tabularx}{\textwidth}{|l|X|}
\hline
\textbf{Name} & EQ\_WINDOW\_IATTR\_HINT\_SWAPSYNC\\
\textbf{Value} & OFF \textbar \ ON // AGL, WGL only\\
\textbf{Default} & \\
\textbf{Description} & \\
\hline
\end{tabularx}\\[\smallskipamount]

\begin{tabularx}{\textwidth}{|l|X|}
\hline
\textbf{Name} & EQ\_WINDOW\_IATTR\_HINT\_DRAWABLE\\
\textbf{Value} & window \textbar \ pbuffer\\
\textbf{Default} & \\
\textbf{Description} & \\
\hline
\end{tabularx}\\[\smallskipamount]

\begin{tabularx}{\textwidth}{|l|X|}
\hline
\textbf{Name} & EQ\_WINDOW\_IATTR\_HINT\_STATISTICS\\
\textbf{Value} & off \textbar \ fastest [on] \textbar \ nicest\\
\textbf{Default} & \\
\textbf{Description} & \\
\hline
\end{tabularx}\\[\smallskipamount]

\begin{tabularx}{\textwidth}{|l|X|}
\hline
\textbf{Name} & EQ\_WINDOW\_IATTR\_PLANES\_COLOR\\
\textbf{Value} & unsigned\\
\textbf{Default} & \\
\textbf{Description} & \\
\hline
\end{tabularx}\\[\smallskipamount]

\begin{tabularx}{\textwidth}{|l|X|}
\hline
\textbf{Name} & EQ\_WINDOW\_IATTR\_PLANES\_ALPHA\\
\textbf{Value} & unsigned\\
\textbf{Default} & \\
\textbf{Description} & \\
\hline
\end{tabularx}\\[\smallskipamount]

\begin{tabularx}{\textwidth}{|l|X|}
\hline
\textbf{Name} & EQ\_WINDOW\_IATTR\_PLANES\_DEPTH\\
\textbf{Value} & unsigned\\
\textbf{Default} & \\
\textbf{Description} & \\
\hline
\end{tabularx}\\[\smallskipamount]

\begin{tabularx}{\textwidth}{|l|X|}
\hline
\textbf{Name} & EQ\_WINDOW\_IATTR\_PLANES\_STENCIL\\
\textbf{Value} & unsigned\\
\textbf{Default} & \\
\textbf{Description} & \\
\hline
\end{tabularx}\\[\smallskipamount]

\begin{tabularx}{\textwidth}{|l|X|}
\hline
\textbf{Name} & EQ\_CHANNEL\_IATTR\_HINT\_STATISTICS\\
\textbf{Value} & OFF \textbar \ FASTEST [ ON ] \textbar \ NICEST\\
\textbf{Default} & \\
\textbf{Description} & \\
\hline
\end{tabularx}\\[\smallskipamount]

\begin{tabularx}{\textwidth}{|l|X|}
\hline
\textbf{Name} & EQ\_COMPOUND\_IATTR\_STEREO\_MODE\\
\textbf{Value} & QUAD \textbar \ ANAGLYPH\\
\textbf{Default} & \\
\textbf{Description} & \\
\hline
\end{tabularx}\\[\smallskipamount]

\begin{tabularx}{\textwidth}{|l|X|}
\hline
\textbf{Name} & EQ\_COMPOUND\_IATTR\_STEREO\_ANAGLYPH\_LEFT\_MASK\\
\textbf{Value} & [ RED GREEN BLUE ]\\
\textbf{Default} & \\
\textbf{Description} & \\
\hline
\end{tabularx}\\[\smallskipamount]

\begin{tabularx}{\textwidth}{|l|X|}
\hline
\textbf{Name} & EQ\_COMPOUND\_IATTR\_STEREO\_ANAGLYPH\_RIGHT\_MASK\\
\textbf{Value} & [ RED GREEN BLUE ]\\
\textbf{Default} & \\
\textbf{Description} & \\
\hline
\end{tabularx}\\[\smallskipamount]

\begin{tabularx}{\textwidth}{|l|X|}
\hline
\textbf{Name} & EQ\_COMPOUND\_IATTR\_UPDATE\_FOV\\
\textbf{Value} & HORIZONTAL [ON] \textbar \ VERTICAL \textbar \ OFF\\
\textbf{Default} & \\
\textbf{Description} & \\
\hline
\end{tabularx}\\[\smallskipamount]
\end{center}

\subsection{Server Section}

{\footnotesize\begin{lstlisting}
server
{
    connection // 0-n times, listening connections of the server
    {
        type       TCPIP | SDP
        TCPIP_port unsigned
        hostname   string
    }

    config {...} // 1-n times, currently only the first one is used by the server
}
\end{lstlisting}}

For an introduction on the concepts of nodes, pipes, windows, channels
and compounds please refer to \sref{sConfig} and \sref{sCompounds}.

\subsubsection{Config Section}

{\footnotesize\begin{lstlisting}
config
{
    latency int    // # of frames nodes may fall behind application
    attributes
    {
         eye_base float // distance between left and right eye
    }

    (node|appNode) {...}// 1-n times, a system in the cluster
                   // 0|1 appNode: launches render thread within app process
    compound {...} // 1-n times
}
\end{lstlisting}}

\subsubsection{Node Section}

{\footnotesize\begin{lstlisting}
{
    name     string
    connection // 0-n times, possible connections to this node
    {
        type          TCPIP | SDP
        TCPIP_port    unsigned
        hostname      string
        command       string      // render client launch command
        command_quote 'character' // launch command argument quote char
        timeout       unsigned    // timeout in milliseconds for launch
    }
    pipe {...} // 1-n times
}
\end{lstlisting}}

\subsubsection{Pipe Section}

{\footnotesize\begin{lstlisting}
pipe
{
    name     string
    port     unsigned     // X server number or ignored
    device   unsigned     // graphics adapter number
    viewport [ viewport ] // default: autodetect
    attributes
    {
        hint_thread   OFF | ON   // default ON
    }

    window {...} // 1-n times
}
\end{lstlisting}}

\subsubsection{Window Section}

{\footnotesize\begin{lstlisting}
window
{
    name     string
    viewport [ viewport ] // wrt pipe, default full screen

    attributes
    {
        hint_stereo         OFF | ON | AUTO
        hint_doublebuffer   OFF | ON | AUTO
        hint_decoration     OFF | ON
        hint_fullscreen     OFF | ON
        hint_swapsync       OFF | ON         // AGL, WGL only
        hint_drawable       window | pbuffer
        hint_statistics     off | fastest [on] | nicest
        planes_color        unsigned
        planes_alpha        unsigned
        planes_depth        unsigned
        planes_stencil      unsigned
    }

    channel {...} // 1-n times
}
\end{lstlisting}}

\subsubsection{Channel Section}

{\footnotesize\begin{lstlisting}
channel
{
    name     string
    viewport [ viewport ] //wrt window, default full window

    attributes
    {
        hint_statistics     OFF | FASTEST [ON] | NICEST
    }
}
\end{lstlisting}}

\subsubsection{Compound Section}

{\footnotesize\begin{lstlisting}
compound
{
    name     string
    channel  string   // where the compound's tasks are executed
    task     [ CLEAR DRAW READBACK ASSEMBLE ] // CULL later
    buffer   [ COLOR DEPTH ]         // default COLOR
    viewport [ viewport ]            // wrt parent compound, sort-first
    range    [ float float ]         // DB-range for sort-last
    pixel    [ int int ]             // pixel decomposition (step size)
    eye      [ CYCLOP LEFT RIGHT ]   // monoscopic or stereo view
    attributes
    {
        stereo_mode                  QUAD | ANAGLYPH    // default QUAD
        // enable visible color components
        stereo_anaglyph_left_mask    [ RED GREEN BLUE ] // default red
        stereo_anaglyph_right_mask   [ RED GREEN BLUE ] // df green blue
        update_FOV                   HORIZONTAL [ON] | VERTICAL | OFF
        assemble                     GPU | CPU          // default GPU
    }

    wall                             // frustum description
    {
        bottom_left  [ float float float ]
        bottom_right [ float float float ]
        top_left     [ float float float ]
    }

    child-compounds

    swapbarrier  // compounds with the same barriername sync swap
    {
        name string
    }
    outputframe
    {
        name string
    }
    inputframe
    {
        name string // corresponding output frame
    }
}
\end{lstlisting}}
